Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems

### 2016-06-13 paper/paper.tex
Better bibtex entry
Fix nonsense connected definition

We introduce definitions and algorithms on undirected and unlabelled graphs; the
extension to general graphs is straightforward and is discussed in \cref{extension}. An undirected
graph $G$ is defined by a finite set of vertices $\operatorname{V}(G)$ and a set of undirected edges
$\operatorname{E}(G) \subseteq \operatorname{V}(G) \times \operatorname{V}(G)$, where $(u, v) \in
\operatorname{E}(G) \Rightarrow (v, u) \in \operatorname{E}(G)$. The neighbourhood of
a vertex $v$, written $\operatorname{N}(G, v)$, is the set of vertices to which it is adjacent, so
$\operatorname{N}(G, v) = \{ u \in \operatorname{V}(G) : (u, v) \in \operatorname{E}(G) \}$.
Given a graph $G$, two vertices $v_s, v_e \in \operatorname{V}(G)$ are connected by a path in $G$ if
there exists a sequence of vertices $(v_0, v_1, \ldots, v_k)$ such that $v_0 = v_s$, $v_k = v_e$,
and each pair of successive vertices is connected by an edge, i.e.\ $\forall i\in \{1 \ldots k \},
(v_{i-1}, v_i) \in \operatorname{E}(G)$. A graph $G$ is \emph{connected} if every non-adjacent pair
of vertices is connected by a path.

@@

In many applications, the common subgraph must be connected
\cite{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,Ehrlich:2011}.
Adding the connectedness requirement makes certain special cases solvable in polynomial time,
including outerplanar graphs of bounded degree \cite{DBLP:journals/algorithms/AkutsuT13} and trees
\cite{DBLP:journals/corr/DroschinskyKM16}, but the general case remains \NP-hard. As illustrated in
\cref{figure:mcsexample}, the MCCS cannot be deduced from the MCS: we need to ensure connectedness
during search. In \cref{mccs-cp}, we show that in CP this may be done in two different ways that
may be combined, and we show in \cref{mccs-cp-eval} that the best results are obtained when
combining them. In \cref{mccs-clique}, we introduce a new way for ensuring connectedness in a
clique-based approach. Finally, we compare CP and our clique-based approach in \cref{mccs-eval}.

>>>>

We introduce definitions and algorithms on undirected and unlabelled graphs; the
extension to general graphs is straightforward and is discussed in \cref{extension}. An undirected
graph $G$ is defined by a finite set of vertices $\operatorname{V}(G)$ and a set of undirected edges
$\operatorname{E}(G) \subseteq \operatorname{V}(G) \times \operatorname{V}(G)$, where $(u, v) \in
\operatorname{E}(G) \Rightarrow (v, u) \in \operatorname{E}(G)$. The neighbourhood of
a vertex $v$, written $\operatorname{N}(G, v)$, is the set of vertices to which it is adjacent, so
$\operatorname{N}(G, v) = \{ u \in \operatorname{V}(G) : (u, v) \in \operatorname{E}(G) \}$.
Given a graph $G$, two vertices $v_s, v_e \in \operatorname{V}(G)$ are connected by a path in $G$ if
there exists a sequence of vertices $(v_0, v_1, \ldots, v_k)$ such that $v_0 = v_s$, $v_k = v_e$,
and each pair of successive vertices is connected by an edge, i.e.\ $\forall i\in \{1 \ldots k \},
(v_{i-1}, v_i) \in \operatorname{E}(G)$. A graph $G$ is \emph{connected} if every distinct pair
of vertices is connected by a path.

@@

In many applications, the common subgraph must be connected
\cite{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,Ehrlich:2011}.
Adding the connectedness requirement makes certain special cases solvable in polynomial time,
including outerplanar graphs of bounded degree \cite{DBLP:journals/algorithms/AkutsuT13} and trees
\cite{droschinsky_et_al:LIPIcs:2016:xxxx}, but the general case remains \NP-hard. As illustrated in
\cref{figure:mcsexample}, the MCCS cannot be deduced from the MCS: we need to ensure connectedness
during search. In \cref{mccs-cp}, we show that in CP this may be done in two different ways that
may be combined, and we show in \cref{mccs-cp-eval} that the best results are obtained when
combining them. In \cref{mccs-clique}, we introduce a new way for ensuring connectedness in a
clique-based approach. Finally, we compare CP and our clique-based approach in \cref{mccs-eval}.
